Corotating Clumps Around Adolescent Low-Mass Stars: Four Years of Complex Quasiperiodic Variables from TESS

### 2023-08-13 paper/ms.tex
merge LR light curve classifxn; update ms

The mosaic in Figure~\ref{fig:cqvs} shows phased light curves for the
\ngoods\ CQVs.  The objects are sorted first in order of the number of
TESS 120-second cadence sectors in which they clearly demonstrated the
CQV phenomenon, and secondarily by descending brightness.  The top
five objects by this metric are TIC~300651846 ($T$$=$13.5, 12
sectors); TIC~402980664 ($T$$=$11.1, 7 sectors); TIC~89463560
($T$$=$13.5, 5 sectors); TIC~363963079 ($T$$=$12.9, 5 sectors); and
TIC~294328887 ($T$$=$14.2, 4 sectors).  The brightest five CQVs span
9.3$<$$T$$<$11.1; the faintest five span 14.5$<$$T$$<$15.0.  The
fastest five have periods spanning 3.6\,hr$<$$P$$<$6.2\,hr, and the
slowest five span 27\,hr$<$$P$$<$38\,hr.  Regarding likely astrometric
or visual binaries, \ngoodhighruwe/\ngoods\ high-quality CQVs have a
Gaia DR2 RUWE$>$2; \nmaybehighruwe/\nmaybes\ of the ambiguous sample
share this characteristic.

@@

A small number of objects at first glance seem reminiscent of
eclipsing binaries, such as TIC~193831684 or TIC~5714469.  We 
believe these cases are unlikely to be eclipsing binaries
due to the additional coherent peaks and troughs in the light curves,
which are
distinct from any binary phenomena of which we are aware.

@@

Of our \ncqvsnodebunked\ confirmed and candidate CQVs, \nnotfieldbanyan\
were associated with a nearby moving group or open cluster using
BANYAN\,$\Sigma$.  The relevant groups are listed in
Table~\ref{tab:thetable}; their ages span $\approx$5-200\,Myr.  The
most prodigious groups were Sco-Cen, Tuc-Hor, and Columba.  Six CQVs
were also identified in the Argus association
\citep{2019ApJ...870...27Z}, which serves as an indirect line of
evidence supporting the reality, and youth, of that group.  The yield
in Sco-Cen is not surprising, since Sco-Cen contains the majority of
pre-main-sequence stars in the solar neighborhood.  However, given the
$\lesssim$$10\%$ completeness of TESS beyond 100\,pc, there are
probably many more CQVs in Sco-Cen that remain to be discovered.

@@

@@

\begin{figure*}[!t]
	\begin{center}
		\centering
		\includegraphics[width=\textwidth]{f6.pdf}
		\vspace{-0.6cm}
		\caption{
      {\bf Evolution of LP 12-502} ($P$=18.6\,h) at fixed period and
      epoch over three years.  Each panel shows one (stacked) TESS
      orbit; small text denotes relative cycle number.  There are 200
      binned black points per cycle.  The TESS pointing law dictates
      the large time gaps between Cycles 64-248, 315-1233, and
      1264-1410; larger gaps tend to yield larger shape changes.  The
      dips usually evolve over tens to hundreds of cycles.  However
      cycles 1233-1264 show a dip that ``switched'' from a depth and
      duration of 3\% and 3\,hr to 0.3\% and 1\,hr over less than one
      cycle (cf.~Figure~\ref{fig:lplc}).
		}
		\label{fig:lp}
	\end{center}
\end{figure*}

\begin{figure*}[!tp]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.48\textwidth]{f7a.pdf}
			\includegraphics[width=0.48\textwidth]{f7b.pdf}
		}
	   \vspace{-0.1cm}

		\subfloat{
			\includegraphics[width=0.48\textwidth]{f7d.pdf}
			\includegraphics[width=0.48\textwidth]{f7c.pdf}
		}
	\end{center}
	\vspace{-0.4cm}
	\caption{
    {\bf River plots of the LP 12-502 light curve}, showing (clockwise
    from top-left) Sectors 18-19, 25-26, 53, and 58-59.  A
    two-harmonic sinusoid has been subtracted to highlight the sharp
    dips.  In Sectors 25-26 (cycles 248-315), periods are visible at
    the fundamental period of 18.5611\,hr, as well as at faster and
    slower relative periods (e.g. $\phi$$\approx$0.3 and
    $\phi$$\approx$0.1).  Similar instances of multiple simultaneous
    periods are also visible in Sector 53 and 58-59.  White chunks
    denote missing data.
	}
	\label{fig:lpriver0}
\end{figure*}

@@

{\sc Dip periods}---Most of the LP 12-502 dips recur with a period of
$P=18.5611 \pm 0.0001$\,hr.  However the river plots
(Figure~\ref{fig:lpriver0}) reveal multiple distinct periodicities in the
light curve for specific dips.  For instance, in sectors 25-26, the
local minimum that develops around cycle 262 has a period faster than
the mean period by $\approx$0.1\%, while some of the trailing local
minima in the main dip complex have periods slower than the mean
period, by $\approx$0.04\%.  In addition to the fundamental period, we
were able to identify at least four distinct periods shown by specific
dips over the full Sectors 18-59 dataset, including periods at
18.5683, 18.5672, 18.5473, and 18.5145\,hr, with a typical measurement
uncertainty of $\approx$0.0002\,hr.  If each period corresponds to an
individual dust or gas clump, then this implies that multiple distinct clumps can orbit
the star simultaneously, at marginally different separations.

@@

		\vspace{-0.4cm}
		\subfloat{
			\includegraphics[width=0.48\textwidth]{f9b.pdf}
		}

		\vspace{-0.4cm}
		\subfloat{
			\includegraphics[width=0.48\textwidth]{f9c.pdf}
		}

@@

\acknowledgments
LGB gratefully acknowledges support from the Heising-Simons
51~Pegasi~b Fellowship.  While many conversations helped encourage
this work, a few thank-yous in particular (from LGB) are owed to
G.~Laughlin, A.~Mann, and J.~Spake for clarifications and ideas.  This
paper relies on data collected by the TESS mission, which are publicly
available from the Mikulski Archive for Space Telescopes (MAST).
Funding for the TESS mission is provided by NASA’s Science Mission
Directorate.  TIC 402980664 in particular was observed at 120-second
cadence thanks to the TESS Guest Investigator programs G022252 (PI:
J.~Schlieder; Sectors 18, 19, 25, 26) and G04168 (PI: R.~Jayaraman;
Sector 53).

{\bf AUTHOR CONTRIBUTION STATEMENTS: PLEASE REVISE IF APPROPRIATE)}
LGB~and RJ conceived the project and executed the dip-based and
Fourier-based analyses, respectively.
SR and RJ vetted the results from the Fourier search.
LGB performed the cluster membership, SED, and variability analyses.
LR contributed an independent SED analysis, and cross-checked the
variability analyses.
AD identified the magnetic B star connection.
LAH contributed to data analysis and interpretation.
G\'AB acquired and maintained the servers used to run the dip-finding
pipeline.
GRR is an architect of the TESS mission.
All authors assisted in manuscript writing and revision.

@@

\startlongtable
\begin{deluxetable}{rrrrrrlrllrrrr}
\tabletypesize{\footnotesize}
%\tabletypesize{\scriptsize}
\tablecaption{Bonafide, candidate, and debunked complex quasiperiodic
  variables from the TESS 2-minute data. {\bf For internal
  review, full versions are available here
  \url{https://www.dropbox.com/scl/fo/twn4s9ckbevf75jqhtoy1/h?rlkey=t5cn8cx2uoc2ptdm9e570kp1b&dl=0}} \label{tab:thetable}}
%\toprule
%TIC & $T$ & $d$ & BP-RP & RUWE & $P$ & assoc & age & teff & rstar & mstar & rcr & qual & Nsectors \\
%\midrule
%\endhead
\startdata
TIC & $T$ & $d$ & $\bprp$ & RUWE & $P$ & Assoc & Age & $T_{\rm eff}$ & $R_\star$ & $M_\star$ & $R$$_{\rm cr}$ & Quality & $N_{\rm sector}$ \\
-- & mag & pc & mag & -- & hr & -- & Myr & K & $R_\odot$ & $M_\odot$ & $R_\star$ & -- & -- \\
\hline
\input{20230613_LGB_RJ_CPV_TABLE_selfnapplied_rounded_data}
\enddata
\tablecomments{This table includes \ngoods\ CQVs (\texttt{Quality}
  flag 1), \nmaybes\ ambiguous CQVs (\texttt{Quality} flag 0), and
  \ndebunked\ impostors (\texttt{Quality} flag -1).  The
  machine-readable version, available online, includes additional
  columns for the Gaia DR2 source identifiers and parameter
  uncertainties.  The age uncertainties are typically $\approx \pm
  10\%$, but can be asymmetric.  {\bf The median temperature, radius, and
  mass uncertainties are $\pm XXX$\,K, $\pm XX$\%, and $\pm YY$\%
  respectively}.  $N_{\rm sector}$ denotes the number of TESS sectors
  for which data are expected to be acquired between July~2018 and
  Oct~2024.   Association names and provenance follow conventions
  adopted by \citet{2018ApJ...856...23G}: 
	ABDMG: AB~Doradus moving group \citep{2015MNRAS.454..593B}.
	ARG: Argus \citep{2019ApJ...870...27Z}.
	%APER: $\alpha$~Persei open cluster \citep{2023AJ....166...14B}.
	BPMG: $\beta$~Pic moving group \citep{2015MNRAS.454..593B}.
  CARN: Carina Near moving group \citep{2006ApJ...649L.115Z}.
	COL: Columba \citep{2015MNRAS.454..593B}.
	EPSC: $\epsilon$ Chamaeleontis \citep{2013MNRAS.435.1325M}.
	LCC: Lower Centaurus Crux \citep{2016MNRAS.461..794P}.
	PERI: Pisces-Eridani \citep{2019AJ....158...77C}.
	PLE: Pleiades \citep{2015ApJ...813..108D}.
	TAU: Taurus \citep{1995ApJS..101..117K}.
	THA: Tucana-Horologium assocation \citep{2015MNRAS.454..593B}.
	TWA: TW Hydrae assocation \citep{2015MNRAS.454..593B}.
	UCL: Upper Centaurus Lupus \citep{2016MNRAS.461..794P}.
	USCO: Upper Scorpius \citep{2016MNRAS.461..794P}.
}
\end{deluxetable}
\clearpage

>>>>

The mosaic in Figure~\ref{fig:cqvs} shows phased light curves for the
\ngoods\ CQVs.  The objects are sorted first in order of the number of
TESS 120-second cadence sectors in which they clearly demonstrated the
CQV phenomenon, and secondarily by descending brightness.  The top
five objects by this metric are TIC~300651846 ($T$$=$13.5, 12
sectors); TIC~402980664 ($T$$=$11.1, 7 sectors); TIC~89463560
($T$$=$13.5, 5 sectors); TIC~363963079 ($T$$=$12.9, 5 sectors); and
TIC~294328887 ($T$$=$14.2, 4 sectors).  The brightest five CQVs span
9.3$<$$T$$<$11.1; the faintest five span 14.5$<$$T$$<$15.0.  The
fastest five have periods spanning 3.6\,hr$<$$P$$<$6.2\,hr, and the
slowest five span 27\,hr$<$$P$$<$38\,hr.  Regarding likely astrometric
or visual binaries, \ngoodhighruwe/\ngoods\ high-quality CQVs have a
Gaia DR2 RUWE$>$2; \nmaybehighruwe/\nmaybes\ of the ambiguous sample
share this characteristic.\footnote{For an explanation of the renormalized unit weight error (RUWE), see the GAIA DPAC technical note \url{http://www.rssd.esa.int/doc_fetch.php? id=3757412}.  Elevated RUWEs imply elevated noise relative to a single-source astrometric model, which can be caused by marginally resolved point sources, astrophysical photometric variability, or astrophysical astrometric motion.}

@@

A small number of objects at first glance seem reminiscent of
eclipsing binaries, such as TIC~193831684, TIC~59836633, or
TIC~5714469.  We believe these cases are unlikely to be eclipsing
binaries due to the additional coherent peaks and troughs in the light
curves, which are distinct from any binary phenomena of which we are
aware.

@@

Of our \ncqvsnodebunked\ confirmed and candidate CQVs, \nnotfieldbanyan\
were associated with a nearby moving group or open cluster using
BANYAN\,$\Sigma$.  The relevant groups are listed in
Table~\ref{tab:thetable}; their ages span $\approx$5-200\,Myr.  The
most prodigious groups were Sco-Cen, Tuc-Hor, and Columba.  Six CQVs
were also identified in the Argus association
\citep{2019ApJ...870...27Z}, which serves as an indirect line of
evidence supporting the reality and youth of that group.  The yield
in Sco-Cen is not surprising, since Sco-Cen contains the majority of
pre-main-sequence stars in the solar neighborhood, any many of its stars
were selected for TESS 120-second cadence observations by various
guest investigators.  Given the
$\lesssim$$10\%$ completeness of TESS beyond 100\,pc, there may
be many more CQVs in Sco-Cen that remain to be discovered.

@@

\begin{figure*}[!t]
	\begin{center}
		\centering
		\includegraphics[width=0.98\textwidth]{f6.pdf}
		\vspace{-0.3cm}
		\caption{
			{\bf Evolution of LP 12-502} ($P$=18.6\,h) at fixed period and
			epoch over three years.  Each panel shows one (stacked) TESS
			orbit; small text denotes relative cycle number.  There are 200
			binned black points per cycle.  The TESS pointing law dictates
			the large time gaps between Cycles 64-248, 315-1233, and
			1264-1410; larger gaps tend to yield larger shape changes.  The
			dips usually evolve over tens to hundreds of cycles.  However
			cycles 1233-1264 show a dip that ``switched'' from a depth and
			duration of 3\% and 3\,hr to 0.3\% and 1\,hr over less than one
			cycle (cf.~Figure~\ref{fig:lplc}).
		}
		\label{fig:lp}
	\end{center}
\end{figure*}

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.46\textwidth]{f7a.pdf}
			\includegraphics[width=0.46\textwidth]{f7b.pdf}
		}
		\vspace{-0.2cm}

		\subfloat{
			\includegraphics[width=0.46\textwidth]{f7d.pdf}
			\includegraphics[width=0.46\textwidth]{f7c.pdf}
		}
	\end{center}
	\vspace{-0.4cm}
	\caption{
		{\bf River plots of the LP 12-502 light curve}, showing (clockwise
		from top-left) Sectors 18-19, 25-26, 53, and 58-59.  A
		two-harmonic sinusoid has been subtracted to highlight the sharp
		dips.  In Sectors 25-26 (cycles 248-315), periods are visible at
		the fundamental period of 18.5611\,hr, as well as at faster and
		slower relative periods (e.g. $\phi$$\approx$0.3 and
		$\phi$$\approx$0.1).  Similar instances of multiple simultaneous
		periods are also visible in Sector 53 and 58-59.  White chunks
		denote missing data.
	}
	\label{fig:lpriver0}
\end{figure*}

@@

@@

{\sc Dip periods}---Most of the LP 12-502 dips repeat with a period of
$P=18.5611 \pm 0.0001$\,hr.  However the river plots
(Figure~\ref{fig:lpriver0}) reveal that the
light curve has specific dips with simultaneous but distinct periods.  For instance, in sectors 25-26, the
local minimum that develops around cycle 262 has a period faster than
the mean period by $\approx$0.1\%, while some of the trailing local
minima in the main dip complex have periods slower than the mean
period, by $\approx$0.04\%.  In addition to the fundamental period, we
were able to identify at least four distinct periods shown by specific
dips over the full Sectors 18-59 dataset, including periods at
18.5683, 18.5672, 18.5473, and 18.5145\,hr, with a typical measurement
uncertainty of $\approx$0.0002\,hr.  If each period corresponds to an
individual dust or gas clump, then this implies that multiple distinct clumps can orbit
the star simultaneously, at marginally different separations.

@@

		\vspace{-0.35cm}
		\subfloat{
			\includegraphics[width=0.48\textwidth]{f9b.pdf}
		}

		\vspace{-0.35cm}
		\subfloat{
			\includegraphics[width=0.48\textwidth]{f9c.pdf}
		}

@@

\acknowledgments
LGB gratefully acknowledges support from the Heising-Simons
51~Pegasi~b Fellowship.  
A few thank-yous in particular (from LGB) are owed to
G.~Laughlin, A.~Mann, and J.~Spake for clarifications and ideas, and to E.~Schaller
and the 51 Peg science team for encouraging lightning talks at the annual summits.
In addition, this paper relied on data collected by the TESS mission, which are publicly
available from the Mikulski Archive for Space Telescopes (MAST).
Funding for the TESS mission is provided by NASA’s Science Mission
Directorate.  TIC 402980664 in particular was observed at 120-second
cadence thanks to the TESS Guest Investigator programs G022252 (PI:
J.~Schlieder; Sectors 18, 19, 25, 26) and G04168 (PI: R.~Jayaraman;
Sector 53).

%{\bf AUTHOR CONTRIBUTION STATEMENTS: PLEASE REVISE IF APPROPRIATE)}
LGB conceived the project and performed the
dip-counting search,
light curve classification,
cluster membership, SED, and variability analyses.
RJ and SR performed the Fourier-based analysis and contributed
to light curve classification.
LR cross-examined the light curve classification, and contributed an independent SED analysis.
AD identified the magnetic B star connection.
LAH contributed to data analysis and interpretation.
G\'AB acquired and maintained the servers used to run the dip-finding pipeline.
GRR is an architect of the TESS mission.
{\bf JNW?  MNG?}
All authors assisted in manuscript writing and revision.

@@

\startlongtable
\begin{deluxetable}{rrrrrrlrllrrrr}
\tabletypesize{\footnotesize}
%\tabletypesize{\scriptsize}
\tablecaption{Bonafide, candidate, and debunked complex quasiperiodic
  variables from the TESS 2-minute data. {\bf For internal
  review, full versions are available here
  \url{https://www.dropbox.com/scl/fo/twn4s9ckbevf75jqhtoy1/h?rlkey=t5cn8cx2uoc2ptdm9e570kp1b&dl=0}} \label{tab:thetable}}
%\toprule
%TIC & $T$ & $d$ & BP-RP & RUWE & $P$ & assoc & age & teff & rstar & mstar & rcr & qual & Nsectors \\
%\midrule
%\endhead
\startdata
TIC & $T$ & $d$ & $\bprp$ & RUWE & $P$ & Assoc & Age & $T_{\rm eff}$ & $R_\star$ & $M_\star$ & $R$$_{\rm cr}$ & Quality & $N_{\rm sector}$ \\
-- & mag & pc & mag & -- & hr & -- & Myr & K & $R_\odot$ & $M_\odot$ & $R_\star$ & -- & -- \\
\hline
\input{20230613_LGB_RJ_CPV_TABLE_selfnapplied_rounded_data}
\enddata
\tablecomments{This table includes \ngoods\ CQVs (\texttt{Quality}
  flag 1), \nmaybes\ ambiguous CQVs (\texttt{Quality} flag 0), and
  \ndebunked\ impostors (\texttt{Quality} flag -1).  The
  machine-readable version, available online, includes additional
  columns for the Gaia DR2 source identifiers and parameter
  uncertainties.  The age uncertainties are typically $\approx \pm
  10\%$, but can be asymmetric.  {\bf The median temperature, radius, and
  mass uncertainties are $\pm XXX$\,K, $\pm XX$\%, and $\pm YY$\%
  respectively}.  $N_{\rm sector}$ denotes the number of TESS sectors
  for which {\it any} data are expected to be acquired between July~2018 and
  Oct~2024.  This number is generally greater than the number of
  sectors for which 120-second cadence data exist.   Association names
  and provenance follow conventions adopted by
  \citet{2018ApJ...856...23G}: 
	ABDMG: AB~Doradus moving group \citep{2015MNRAS.454..593B}.
	ARG: Argus \citep{2019ApJ...870...27Z}.
	%APER: $\alpha$~Persei open cluster \citep{2023AJ....166...14B}.
	BPMG: $\beta$~Pic moving group \citep{2015MNRAS.454..593B}.
  CARN: Carina Near moving group \citep{2006ApJ...649L.115Z}.
	COL: Columba \citep{2015MNRAS.454..593B}.
	EPSC: $\epsilon$ Chamaeleontis \citep{2013MNRAS.435.1325M}.
	LCC: Lower Centaurus Crux \citep{2016MNRAS.461..794P}.
	PERI: Pisces-Eridani \citep{2019AJ....158...77C}.
	PLE: Pleiades \citep{2015ApJ...813..108D}.
	TAU: Taurus \citep{1995ApJS..101..117K}.
	THA: Tucana-Horologium assocation \citep{2015MNRAS.454..593B}.
	TWA: TW Hydrae assocation \citep{2015MNRAS.454..593B}.
	UCL: Upper Centaurus Lupus \citep{2016MNRAS.461..794P}.
	USCO: Upper Scorpius \citep{2016MNRAS.461..794P}.
}
\end{deluxetable}
\clearpage
